The Surprising Runtime Fragility Found when Querying a Single Table

### 2015-05-12 Documentation.tex
!

\documentclass[11pt]{scrartcl}
\usepackage{PDBF}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage[a4paper,vmargin={28mm,28mm},hmargin={28mm,28mm}]{geometry}

\title{PDBF Documentation}
\author{
 Patrick Bender\\
  \texttt{s9pabend@stud.uni-saarland.de}
  \and
 Christian Goldstein\\
  \texttt{c.goldstein@goldstein-systemberatung.de}
}
\date{\today{}, Saarbrücken}

@@

TODO: Write small abstract \newpage

\section{Chart options}
Macros: \\[3pt]
\verb|\lineChart[options]{query}| \\
\verb|\barChart[options]{query}| \\[8pt]
\noindent Options: \\[3pt]
\begin{tabular}{ll}
\textbf{width} & \\
Desc.:	&	Sets the width of the chart. \\
Argument:	&LaTeX length \\
Default:&	no default value. value must always be set! \\[4pt]

@@

\textbf{options} \\
Desc.:	&	Sets options that are directly passed to the dygraphs library (\href{http://dygraphs.com/options.html}{$\rightarrow$ Documentation}). \\
& You need to wrap the JSON-String with \verb|{}| if you want to use \verb|[| or \verb|]| or \verb|,|. \\
Argument: 	&JSON-String \\
Default:&	\verb|{}| (which means empty object) \\[4pt]

\textbf{logscale} \\
Desc.:	&	If set, the y axis uses log scale. \\
& Can also be redefined globally (\verb|\pdbfLogscale|) \\
Argument:&	\verb|{true, false}| \\
Default:&	\verb|false| \\[4pt]

@@

\textbf{legendpos} \\
Desc.: 	&	Sets the position of the legend of the chart. \\
& Can also be redefined globally (\verb|\pdbfLegendpos|).\\
Argument:&	\verb|{topright, topcenter, topleft, belowleft, belowcenter, belowright}| \\
Default:&	\verb|topright| \\[4pt]
\end{tabular} \\

\newpage
\section{F.A.Q.}
\begin{itemize}
\item F.: The chart is horribly slow when i try to zoom! \\
A.: Add the following to options: animatedZooms: false 
\end{itemize}

>>>>

\documentclass[11pt]{scrartcl}
\usepackage{PDBF}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage{ngerman}
\usepackage[a4paper,vmargin={20mm,20mm},hmargin={28mm,28mm}]{geometry}

\title{PDBF Documentation}
\author{
 Patrick Bender\\
  \texttt{s9pabend@stud.uni-saarland.de}
}
\date{\today{}, Saarbr"ucken}

@@

%TODO: Write small abstract

% TODO: picasso diag

%Required packages:
%\RequirePackage{zref-savepos}
%\RequirePackage{zref-abspage}
%\RequirePackage{zref-user}
%\RequirePackage{xcolor}
%\RequirePackage{graphicx}
%\RequirePackage{xstring}
%\RequirePackage{xparse}
%\RequirePackage{geometry}

%need to use geometry package

\newpage

\section{Chart options}
Macros: \\[3pt]
\verb|\chart[options][queryForOverlay]{queryForPage}| \\[8pt]
\noindent Options: \\[3pt]
\begin{tabular}{ll}
\textbf{chartType} & \\
Desc.:	&	Sets the type of the chart. \\
Argument:	&\verb|line or bar or signatureplot| (more info here)  \\
Default:&	\verb|line| \\[4pt]

\textbf{width} & \\
Desc.:	&	Sets the width of the chart. \\
Argument:	&LaTeX length \\
Default:&	no default value. value must always be set! \\[4pt]

@@

\textbf{options} \\
Desc.:	&	Sets options that are directly passed to the c3 chart library (\href{http://c3js.org/reference.html}{$\rightarrow$ Documentation}). \\
& You need to wrap the JSON-String with \verb|{}| if you want to use \verb|[| or \verb|]| or \verb|,|. \\
Argument: 	&JSON-String \\
Default:&	\verb|{}| (which means empty object) \\[4pt]

%TODO: not yet implemented
%\textbf{logscale} \\
%Desc.:	&	If set, the y axis uses log scale. \\
%& Can also be redefined globally (\verb|\pdbfLogscale|) \\
%Argument:&	\verb|true or false| \\
%Default:&	\verb|false| \\[4pt]

@@

%TODO: not yet implemented
%\textbf{legendpos} \\
%Desc.: 	&	Sets the position of the legend of the chart. \\
%& Can also be redefined globally (\verb|\pdbfLegendpos|).\\
%Argument:&	\verb|{topright, topcenter, topleft, belowleft, belowcenter, belowright}| \\
%Default:&	\verb|topright| \\[4pt]
\end{tabular} \\

\newpage
\section{F.A.Q.}
\begin{itemize}
\item Q.: The overlay is not on the right position \\
A.: Most likely you use pages with different sizes in your document. This is currently not supported by PBDF 
\item Q.: I get \verb|Error: paperwidth value missing! Did you forgot to specify the| \\ \verb|papersize via the geometry package?| when compiling my document \\
A.: You need to explicitly specify the papersize of your document via the geometry package (e.g. \verb|\usepackage[letterpaper]{geometry}|)
\end{itemize}

### 2015-05-12 test.tex
!

\begin{document}
\maketitle
\begin{wrapfigure}{r}{0.47\textwidth}
{ \centering %
\chart[width=0.4\textwidth, height=0.3\textwidth, chartType=Line, options={"title": "Title"}]{SELECT * FROM data1;}%
\\ }
\vspace{-5pt}
\caption{Graph1}
\vspace{-20pt}
\end{wrapfigure} ~\\
Test Queries:\\[3pt]
\sql{SELECT * FROM test;}\\[3pt]
\sql{SELECT col2 FROM test WHERE col1=1;}\\[3pt]
\sqlmult{INSERT INTO test}{INSERT INTO test VALUES (1, 123);}\\
\sqlmult{VALUES (1, 123);}{INSERT INTO test VALUES (1, 123);}\\[8pt]
Lorem ipsum \sql{SELECT * FROM test;} dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.\\

@@

@@

>>>>

\begin{document}
\maketitle
\begin{wrapfigure}{r}{0.47\textwidth}
{ \centering %
\chart[width=0.4\textwidth, height=0.3\textwidth, chartType=Line, options={"title": "Title"}]{SELECT * FROM data1;}%
\\ }
\vspace{-5pt}
\caption{Graph1}
\vspace{-20pt}
\end{wrapfigure} ~\\
Test Queries:\\[3pt]
\sql{SELECT * FROM test;}\\[3pt]
\sql{SELECT col2 FROM test WHERE col1=1;}\\[3pt]
\sql[INSERT INTO test]{INSERT INTO test VALUES (1, 123);}\\
\sql[VALUES (1, 123);]{INSERT INTO test VALUES (1, 123);}\\[8pt]
Lorem ipsum \sql{SELECT * FROM test;} dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.\\

@@

[SELECT

@@

    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;]

### 2015-05-12 testdata/chunked_column_layouts/draft.tex
!

\begin{table}[!b] 
\pivotTable[width=0.478\textwidth, height=0.7478\textwidth, scale=0.666, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time, aggregationBig=Minimum, aggregationattributeBig=Query time (in Sec)]
{SELECT
m.cpuarchitecture AS Machine,
IIF(d.tableschema = '(long a, long b)', 'long', IIF(d.tableschema = '(char a, char b)', 'char', IIF(d.tableschema = '(int a, int b)', 'int', 'n/a')))   AS [Table schema],
IIF(q.querytext = 'SELECT MIN(a+b) FROM t', 'Q1', 'Q2')     AS [Query],
c.compiler      AS Compiler,
c.optlevel      AS [Opt. level],
l.name          AS [Layout name],
CAST(COALESCE(IIF(i.templateparam, 'C', 'R'), 'n/a') AS VARCHAR) AS [Chunk size provided at],
ROUND(v.runtime, 3) 		AS [Query time]
FROM fact_experiment AS v
    JOIN dim_machine m ON m.id = v.machineid
    JOIN dim_dataset d ON d.id = v.datasetid
    JOIN dim_compiler c ON c.id = v.compilerid
    JOIN dim_layout l ON l.id = v.layoutid
    JOIN dim_implementation i ON i.id = v.implementationid
    JOIN dim_query q ON q.id = v.queryid
ORDER BY m.id DESC, d.tableschema, q.querytext,
    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
{SELECT
m.cpu + ' (' +  m.cpuarchitecture + ')' AS Machine,
d.tableschema   AS [Table schema],
q.querytext     AS [Query],
c.compiler      AS Compiler,
c.optlevel      AS [Opt. level],
l.name          AS [Layout name],
CAST(COALESCE(l.chunksize, 'n/a') AS VARCHAR) AS [Chunk size (tuples)],
IIF(i.templateparam, 'Compile time', 'Runtime') AS [Chunk size provided at],
IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') AS [Chunk sizes considered],
t.timestamp 	AS [Time of experiment],
ROUND(v.runtime, 3) 		AS [Query time (in Sec)]
FROM fact_experiment AS v
    JOIN dim_machine m ON m.id = v.machineid
    JOIN dim_dataset d ON d.id = v.datasetid
    JOIN dim_compiler c ON c.id = v.compilerid
    JOIN dim_layout l ON l.id = v.layoutid
    JOIN dim_implementation i ON i.id = v.implementationid
    JOIN dim_query q ON q.id = v.queryid
    JOIN dim_time t ON t.id = v.timeid
ORDER BY m.id DESC, d.tableschema, q.querytext,
    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
\vspace*{-5pt}
\caption{The best layouts and most efficient ways of implementing a given query on a given table and executed on a given machine.}  
\label{fig:problem_bests_table}
\vspace{-10pt}
\end{table}

>>>>

\begin{table}[!b] 
\pivotTable[width=0.478\textwidth, height=0.7478\textwidth, scale=0.666, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time]
[SELECT 
  m.cpu + ' (' + m.cpuarchitecture + ')' 						AS Machine, 
  d.tableschema 											AS [Table schema], 
  q.querytext 												AS [Query], 
  c.compiler 												AS Compiler, 
  c.optlevel 												AS [Opt. level], 
  l.name 													AS [Layout name], 
  CAST(COALESCE(l.chunksize, 'n/a') AS VARCHAR) 			AS [Chunk size (tuples)], 
  IIF(i.templateparam, 'Compile time', 'Runtime') 				AS [Chunk size provided at], 
  IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') 			AS [Chunk sizes considered], 
  runtimes 												AS [Query time] 
  FROM (SELECT machineid, datasetid, queryid, compilerid, layoutid, implementationid, GRUBBS_FILTER(ARRAY(runtime)) as runtimes FROM fact_experiment GROUP BY machineid, datasetid, queryid, compilerid, layoutid, implementationid) AS v 
  JOIN dim_machine m ON m.id = v.machineid 
  JOIN dim_dataset d ON d.id = v.datasetid 
  JOIN dim_compiler c ON c.id = v.compilerid 
  JOIN dim_layout l ON l.id = v.layoutid 
  JOIN dim_implementation i ON i.id = v.implementationid 
  JOIN dim_query q ON q.id = v.queryid 
  ORDER BY m.id DESC, d.tableschema, q.querytext, c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;]
{SELECT
m.cpuarchitecture 																													AS Machine,
IIF(d.tableschema = '(long a, long b)', 'long', IIF(d.tableschema = '(char a, char b)', 'char', IIF(d.tableschema = '(int a, int b)', 'int', 'n/a')))   			AS [Table schema],
IIF(q.querytext = 'SELECT MIN(a+b) FROM t', 'Q1', 'Q2')     																				AS [Query],
c.compiler      																														AS Compiler,
c.optlevel      																														AS [Opt. level],
l.name          																														AS [Layout name],
CAST(COALESCE(IIF(i.templateparam, 'C', 'R'), 'n/a') AS VARCHAR) 																		AS [Chunk size provided at],
runtimes																																AS [Query time]
  FROM (SELECT machineid, datasetid, queryid, compilerid, layoutid, implementationid, GRUBBS_FILTER(ARRAY(runtime)) as runtimes FROM fact_experiment GROUP BY machineid, datasetid, queryid, compilerid, layoutid, implementationid) AS v 
  JOIN dim_machine m ON m.id = v.machineid 
  JOIN dim_dataset d ON d.id = v.datasetid 
  JOIN dim_compiler c ON c.id = v.compilerid 
  JOIN dim_layout l ON l.id = v.layoutid 
  JOIN dim_implementation i ON i.id = v.implementationid 
  JOIN dim_query q ON q.id = v.queryid 
  ORDER BY m.id DESC, d.tableschema, q.querytext, c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
\vspace*{-5pt}
\caption{The best layouts and most efficient ways of implementing a given query on a given table and executed on a given machine.}  
\label{fig:problem_bests_table}
\vspace{-10pt}
\end{table}
